Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
State of the art has been reviewed

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a
``language for the internet'', providing some advantages such as multi-architecture compatibility or
security mechanisms, and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using compiled Java
programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers. 
%% We need to only reference this section?:
%%------------------- BEGIN
% This paper % (Paloma) No really need for that "also", plus it complicates reading
% establishes the desirable features of
% any {\em global computing} infrastructure, that can also be applied to
% distributed evolutionary computation experiments that rely on
% volunteer computing: scalability, heterogeneity, fault tolerance,
% adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
% reasonable performance. But hierarchy, this is, the capability of
% preferentially using resources in the same organization, is a feature
% that could be either disregarded or included within adaptivity, which
% leaves us with eight desirable properties. 
% % (Paloma) Re-write. For example, "But is hierarchy, this is, the
% % capability of preferentially using resources in the same
% % organization, the one that could be either disregarded or included
% % in adaptivity, which leaves us with eight desirable properties." 
% % (Paloma) And I re-wrote what I proposed yesterday.
% % thanks :-) -JJ
% %%------------------- END
% %% EDIT:
% %%------------------- BEGIN
% The fulfillment of these properties could be achieved by using any of the
% languages and associated technologies mentioned above, provided that the user had
% installed the required virtual machine plugin first. 
% %%------------------- END
The fulfillment of the desirable features of any distributed evolutionary 
computation experiments that rely on volunteer computing \cite{2016arXiv160101607Manom} could be achieved by using any of the
languages and associated technologies, mentioned above, provided that the user had
installed the required virtual machine plugin first.

@@

JavaScript can be used either for unwitting, that is, stealth,
\cite{unwitting-ec} or volunteer 
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210}

@@

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts. There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}. This is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or by farming out evaluations to all
available nodes. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\cite{nogueras2015studying} or be included in a broader study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}.

On the other hand, measuring the performance of the resulting metacomputer
involves understanding its dynamics. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
%But the raw material of           
%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
But some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal, and
Gamma distributions 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}. Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. 
%Mario: erase: Next we describe the framework itself.
As far as we know, this paper presents one of the few experimental
results using a volunteer computing system which uses  % (Paloma) Change to "which" in order to not repeating "that" all the time
very diverse computational resources; these will be described next.

>>>>

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
 including Flash animations and VBScript, ActiveX, or Java applets were
used with the same purpose; for instance, Java was pointed out in \cite{soares1998get} as a
``language for the internet'', providing some advantages such as multi-architecture compatibility or
security mechanisms, and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using compiled Java
programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers. 
%% We need to only reference this section?:
%%------------------- BEGIN
% This paper % (Paloma) No really need for that "also", plus it complicates reading
% establishes the desirable features of
% any {\em global computing} infrastructure, that can also be applied to
% distributed evolutionary computation experiments that rely on
% volunteer computing: scalability, heterogeneity, fault tolerance,
% adaptive parallelism, volunteer's safety, anonymity, hierarchy, ease of use, and
% reasonable performance. But hierarchy, this is, the capability of
% preferentially using resources in the same organization, is a feature
% that could be either disregarded or included within adaptivity, which
% leaves us with eight desirable properties. 
% % (Paloma) Re-write. For example, "But is hierarchy, this is, the
% % capability of preferentially using resources in the same
% % organization, the one that could be either disregarded or included
% % in adaptivity, which leaves us with eight desirable properties." 
% % (Paloma) And I re-wrote what I proposed yesterday.
% % thanks :-) -JJ
% %%------------------- END
% %% EDIT:
% %%------------------- BEGIN
% The fulfillment of these properties could be achieved by using any of the
% languages and associated technologies mentioned above, provided that the user had
% installed the required virtual machine plugin first. 
% %%------------------- END
The fulfillment of the desirable features of any distributed evolutionary 
computation experiments that rely on volunteer computing \cite{2016arXiv160101607Manom} could be achieved by using any of the
languages and associated technologies mentioned above, provided that the user had
installed the required virtual machine plugin first.

@@

JavaScript can be used either for unwitting, that is, stealth,
\cite{unwitting-ec} or volunteer 
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210}.

@@

Additionally, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts. There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}. This is actually not such a big problem in
algorithms such as the Evolutionary Algorithm (EA) that can easily be
parallelized via population splitting or by farming out evaluations to all
available nodes. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
the more general context of distributed computing 
\cite{nogueras2015studying} or be included in a broader study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}.

Moreover, measuring the performance of the resulting metacomputer
involves understanding its dynamics. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
%But the raw material of           
%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
But some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal, and
Gamma distributions 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}. % (Paloma) I have a doubt here, should this reference be anonymised or not?
Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. 
%Mario: erase: Next we describe the framework itself.
As far as we know, this paper presents one of the few experimental
results using a volunteer computing system which uses  % (Paloma) Change to "which" in order to not repeating "that" all the time
very diverse computational resources; these will be described next.
